The Units of Permissionless Consensus: \\ Towards Mobile and Edge Computing

### 2022-10-27 theoretical-consensus/article/article.tex
add refs

Following the line of the previous sections, it has been reasoned about the need 
for permissionless consensus when there are already well known and established consensus protocols
that work in trusted environments \cite{castro1999practical, miller2016honey}. However, even those protocols have
their own limitations, not only in terms of trust, fault-tolerance, centrality, permissions,
or bottlenecks, but also in terms of scalability \cite{miller2016honey}. Trying to put some effort on decentralization, 
Byzantine-Fault-Tolerant consensus protocols are not known for their performance when the network grows
in size. The more participants there are, the more messages need to be exchanged
between them, and the more time it takes to reach consensus, even if assuring deterministic finality \cite{decker2016bitcoin}. 
This is a problem that is not only related to the number of participants, but also to the
communication fashion and bandwidth, and the computational capacity of the devices that
participate in the consensus protocol. 
Summing up, the need for permissionless consensus is then justified by the fact that
permissioned protocols are not compatible with the requirements of the new generation of
distributed systems, especially in the context of blockchain networks. 
These requirements include dealing with today's sparse networks of anonymously and dynamically
participating devices, without interrupting consensus and while battling Sybil attacks \cite{survey-dist-consensus}.
Fundamentally, the permissionless consensus problem is the need for a consensus protocol that
can be run in a distributed and decentralized environment, where the participants are unknown and untrusted,
and where the network is bigger, sparser and unpredictably less reliable.

@@

Technically, permissionless environments allow for larger networks
that depict lower connectivity between the participants. Operationally,
everything is expected to happen in an asynchronous or partially synchronous fashion, 
and the number of transactions is expected to be smaller than the 
permissioned counterparts. Nevertheless, participation is free, and the
governance is not centralized, but rather distributed and public. 
The identity of the participants is secured or semi-secure as it often relies on
pseudonymity for protecting the nodes' identity, while enabling full transparency 
in regard to the rest of the network's content and operation \cite{xiao2019distributed}.
The goal of permissionless consensus, as generally for any consensus protocol, is to reach
agreement on a single value, or a set of values. However, due to the nature of the
protocols, the values that are agreed upon end up establishing the serialization of the
transactions that are executed in the network, and so establishing time consciousness and 
total order of the events.

As pointed out by \cite{xiao2019distributed} and then referenced by \cite{survey-dist-consensus}, 
similarly to the permissioned counterparts, permissionless consensus protocols 
aim at achieving the following properties: Termination, Agreement, Validity, 
and Integrity. Without going into a lot of details, of these properties, 
Agreement and Integrity are the most important ones, as they are the ones that
guarantee the correctness of the consensus protocols. Termination and Validity
are generally related to any classic consensus problem.

@@

In the classical Nakamoto consensus protocol, the generation of a block, to be proposed
for further network agreement, complies with the unit of computational work needed to
create, or rather find, a verifiable proof of the effort spent on assembling the block \cite{nakamoto2008bitcoin}. 
This essentially requires brute forcing the search for a cryptographic hash value for the
aggregation of the block information with a nonce, such that this hash value satisfies
a difficulty threshold (See Algorithm \ref{alg:BlockGeneration}), which gets adjusted dynamically over time, to maintain the network overall 
requirement for the block generation interval \cite{survey-dist-consensus}.

@@

Despite the harmony of all the aforementioned clever characteristics,
the proof-of-work mechanism has some well known drawbacks. 
First, the block generation is computationally expensive, and so, the
energy consumption associated with it is also high. This is caused by the
dynamically adjusted difficulty threshold, which is a function of the block 
generation interval. For example, in the Bitcoin network, to maintain the 10 
minutes interval, the hashing difficulty
increases with the increase of the hashing power, and so too the energy needed for computing that proof. 
However, reducing the block generation interval would increase the
probability of forks, which results in a more frequent waste of useful work,
loosening the security of the network. Increasing the block size would also
have the same security effect due to the increased propagation delay.

>>>>

Following the line of the previous sections, it has been reasoned about the need 
for permissionless consensus when there are already well known and established consensus protocols
that work in trusted environments \cite{castro1999practical, miller2016honey}. However, even those protocols have
their own limitations, not only in terms of trust, fault-tolerance, centrality, permissions,
or bottlenecks, but also in terms of scalability \cite{miller2016honey}. Trying to put some effort on decentralization, 
Byzantine-Fault-Tolerant consensus protocols are not known for their performance when the network grows
in size. The more participants there are, the more messages need to be exchanged
between them, and the more time it takes to reach consensus, even if assuring deterministic finality \cite{decker2016bitcoin}. 
This is a problem that is not only related to the number of participants, but also to the
communication fashion and bandwidth, and the computational capacity of the devices that
participate in the consensus protocol. 
Summing up, the need for permissionless consensus is then justified by the fact that
permissioned protocols are not compatible with the requirements of the new generation of
distributed systems, especially in the context of blockchain networks. 
These requirements include dealing with today's sparse networks of anonymously and dynamically
participating devices, without interrupting consensus and while battling Sybil attacks \cite{8629877, survey-dist-consensus}.
Fundamentally, the permissionless consensus problem is the need for a consensus protocol that
can be run in a distributed and decentralized environment, where the participants are unknown and untrusted,
and where the network is bigger, sparser and unpredictably less reliable.

@@

Technically, permissionless environments allow for larger networks
that depict lower connectivity between the participants. Operationally,
everything is expected to happen in an asynchronous or partially synchronous fashion, 
and the number of transactions is expected to be smaller than the 
permissioned counterparts. Nevertheless, participation is free, and the
governance is not centralized, but rather distributed and public. 
The identity of the participants is secured or semi-secure as it often relies on
pseudonymity for protecting the nodes' identity, while enabling full transparency 
in regard to the rest of the network's content and operation \cite{xiao2019distributed}.
The goal of permissionless consensus, as generally for any consensus protocol, is to reach
agreement on a single value, or a set of values. However, due to the nature of the
protocols, the values that are agreed upon end up establishing the serialization of the
transactions that are executed in the network, and so establishing time consciousness and 
total order of the events \cite{8629877}.

As pointed out by \cite{xiao2019distributed, 8629877} and then referenced by \cite{survey-dist-consensus}, 
similarly to the permissioned counterparts, permissionless consensus protocols 
aim at achieving the following properties: Termination, Agreement, Validity, 
and Integrity. Without going into a lot of details, of these properties, 
Agreement and Integrity are the most important ones, as they are the ones that
guarantee the correctness of the consensus protocols. Termination and Validity
are generally related to any classic consensus problem.

@@

In the classical Nakamoto consensus protocol, the generation of a block, to be proposed
for further network agreement, complies with the unit of computational work needed to
create, or rather find, a verifiable proof of the effort spent on assembling the block \cite{nakamoto2008bitcoin}. 
This essentially requires brute forcing the search for a cryptographic hash value for the
aggregation of the block information with a nonce, such that this hash value satisfies
a difficulty threshold (See Algorithm \ref{alg:BlockGeneration}), which gets adjusted dynamically over time, to maintain the network overall 
requirement for the block generation interval \cite{8629877, survey-dist-consensus}.

@@

Despite the harmony of all the aforementioned clever characteristics,
the proof-of-work mechanism has some well known drawbacks. 
First, the block generation is computationally expensive, and so, the
energy consumption associated with it is also high. This is caused by the
dynamically adjusted difficulty threshold, which is a function of the block 
generation interval. For example, in the Bitcoin network, to maintain the 10 
minutes interval, the hashing difficulty
increases with the increase of the hashing power, and so too the energy needed for computing that proof. 
However, reducing the block generation interval would increase the
probability of forks, which results in a more frequent waste of useful work,
loosening the security of the network. Increasing the block size would also
have the same security effect due to the increased propagation delay \cite{8629877}.
